---
title: "Lab 03 - Nobel laureates"
author: Lindsey Wilson
date: 01/26/2023
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

The `nobel` dataset has 935 observations and 26 variables:

```{r counts}
# this gives the number of observations we have
nobel %>%
  summarize(n())

#and this tells us the names/numbers of our variables
names(nobel)
```

### Exercise 2

The code below creates a new data frame called `nobel_living` that filters the `nobel` dataset for living people.
If we did this rght, we should be left with a 228 observations:

```{r nobel-living}
nobel_living <- nobel %>%
  filter(!is.na(country),
         gender != "org",
         is.na(died_date)
         )

count <- nobel_living %>%
  summarize(n())

message("the nobel_living dataset contains ", count, " observations")
```

### Exercise 3

Now we want to visualize how likely US vs. non-US scientists are to win Nobel prizes in the sciences.
To do this. we first have to add a variable to the `nobel_living` data frame called `country_us`.
This variable takes the value of "USA" if the person is from the US, and "other" if not:

```{r mutate}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

Next we have to make a new data frame called `nobel_living_science` that only considers prizes in physics, medicine, chemistry, and economics:

```{r}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

And finally, we can use this new data frame to make a visualization that compares the number of Nobel laureates from each science category that are US-based vs. non-US-based.
From this we can see that US-based researchers are much more likely to win Nobel prizes, especially in economics.

```{r}
ggplot(data = nobel_living_science,
       mapping = aes(x = country_us,
                     fill = country_us)) + 
  geom_histogram(stat = "count") +
  coord_flip() +
  facet_wrap( ~ category, ncol = 2)
```

### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
