---
title: "Lab 03 - Nobel laureates"
author: Lindsey Wilson
date: 01/26/2023
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

The `nobel` dataset has 935 observations and 26 variables:

```{r counts}
# this gives the number of observations we have
nobel %>%
  summarize(n())

#and this tells us the names/numbers of our variables
names(nobel)
```

### Exercise 2

The code below creates a new data frame called `nobel_living` that filters the `nobel` dataset for living people.
If we did this rght, we should be left with a 228 observations:

```{r nobel-living}
nobel_living <- nobel %>%
  filter(!is.na(country),
         gender != "org",
         is.na(died_date)
         )

count <- nobel_living %>%
  summarize(n())

message("the nobel_living dataset contains ", count, " observations")
```

### Exercise 3

Now we want to visualize how likely US vs. non-US scientists are to win Nobel prizes in the sciences.
To do this. we first have to add a variable to the `nobel_living` data frame called `country_us`.
This variable takes the value of "USA" if the person is from the US, and "Other" if not:

```{r mutate}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

Next we have to make a new data frame called `nobel_living_science` that only considers prizes in physics, medicine, chemistry, and economics:

```{r filter-science}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

And finally, we can use this new data frame to make a visualization that compares the number of Nobel laureates from each science category that are US-based vs. non-US-based.
From this we can see that US-based researchers are much more likely to win Nobel prizes, especially in economics.

```{r US-based-plot}
ggplot(data = nobel_living_science,
       mapping = aes(x = country_us,
                     fill = country_us)) + 
  geom_histogram(stat = "count") +
  coord_flip() +
  facet_wrap( ~ category, ncol = 2)
```

### Exercise 4

So if being *based* in the US increases your chances of winning a Nobel prize in the sciences, does it matter whether or not you were *born* in the US?
Let's find out.
Just like last time, we start by adding a variable to `nobel_living_science` called `born_country_us`.
This variable takes the value of "USA" if the researcher was born in the US and "Other" if not:

```{r born-US}
nobel_living_science <- nobel_living_science %>%
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
    )
```

As a fun aside, we can count how many individuals in our modified dataset were born in the US:

```{r born-US-count}
born_US_count <- nobel_living_science %>%
  filter(born_country_us == "USA") %>%
  summarize(n())

message(born_US_count, " individuals in the the nobel_living_science dataset were born in the US")
```

### Exercise 5

Finally, we can adapt the code. from Ex. 3 to see how many of the US/foreign based winners were born in the US vs. outside the US:

```{r US-born-plot}
ggplot(data = nobel_living_science,
       mapping = aes(x = country_us,
                     fill = born_country_us)) + 
  geom_histogram(stat = "count") +
  coord_flip() +
  facet_wrap( ~ category, ncol = 2)
```

This allows us to see that, while a majority of the US-based winners in the sciences were also US-born, a sizeable chunk were born outside the US.
Therefore, it looks like being *in* the US helps for winning a prize, but being *from* the US isn't as strong a predictor.
This supports the conclusions drawn by the Buzzfeed article, since the only way you could be based in the US but not from there is if you immigrated.

### Exercise 6

So, where are those immigrant Nobel laureates from? 
The code below allows us to answer that question by filtering from `nobel_living_sciences` those winners who were based in the US at the time of their award  but who were born outside of it.
We can then generate a frequency plot that tells us the names and counts of the countries where these individuals were born:

```{r born-country-counts}
born_country_counts <- nobel_living_science %>%
  filter(country_us == "USA",
         born_country_us == "Other") %>%
  count(born_country, sort = TRUE)

print(born_country_counts, n = 100)
```

