---
title: "Lab 03 - Nobel laureates"
author: Lindsey Wilson
date: 01/26/2023
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

The `nobel` dataset has 935 observations and 26 variables:

```{r counts}
# this gives the number of observations we have
nobel %>%
  summarize(n())

#and this tells us the names/numbers of our variables
names(nobel)
```

We want to filter this for living people.
The code to do that is below.
If we did this right, we. should have 228 observations:

```{r}
nobel_living <- nobel %>%
  filter(!is.na(country),
         gender != "org",
         is.na(died_date)
         )

nobel_living %>%
  summarize(n())
```
### Exercise 2

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
